Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-11-21 Ben_Manuscripts/structure_paper/response_to_reviewers.tex

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{gensymb}

\externaldocument[S-]{supporting_information}
\externaldocument[M-]{Final_Draft}

\begin{document}

%MRS: making the title a bit smaller, removing date.
\begin{center}
\textbf{Response to reviewers: Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes}: \\
Authors: Benjamin J. Coscia, Joseph Yelk, Matthew A. Glaser, Douglas L. Gin, Xunda Feng, and Michael R. Shirts
\end{center}

@@

    \end{quote}

    Author reply: We thank the reviewer for offering their interpretation of
    the purpose of our study, as it is important that we properly convey our
    message. The main purpose of our work is to give a detailed description of the
    structure of a lyotropic liquid crystal membrane backed by experimental data.
    This work lays a foundation for future transport studies. By matching 2D-WAXS
    data, we have attempted to unveil the most experimentally-consistent structure
    possible. To clarify this purpose, we have modified and added text to the
    paragraph on page~\pageref{M-rework:purpose} of the introduction, in which we
    state our perceived importance of knowing the most experimentally-consistent
    structure:

    Old text: ``A molecular-level understanding of LLC polymer membrane
    structure, enabled by molecular dynamics (MD) simulations, can provide
    guidelines to reduce the large chemical space available to design monomers for
    creation of separation-specific membranes. Useful molecular-level modeling
    should incorporate a detailed picture of the nanoscopic pore structure, which
    is crucial to understanding the role of monomer structure in solute transport
    and membrane design.

    New text: ``A molecular-level understanding of LLC polymer membrane
    structure, enabled by molecular dynamics (MD) simulations, will enhance our
    understanding of the role of monomer structure in solute transport and provide
    guidelines to reduce the large chemical space available to design monomers for
    creation of separation-specific membranes. It is important that we model the
    system with a nanoscopic pore structure that is maximally consistent with
    experimental data in order to best approximate the chemical environment experienced
    by solutes in future simulated transport studies."

@@

%This is unlikely to be achieved if one assembled these systems using
%    united atom or coarse grained models due to the large number of potential
%    metastable configurational basins which plague a system like ours which exhibts
%    such slow dynamics.

    We do not intend for this study to be a validation of GAFF for this
    particular system. We believe we would come to similar conclusions using other
    force fields, however validation of this hypothesis is left for a future study.
    We chose GAFF because it is widely used to simulate organic molecules and there
    are no catch-all force fields built for liquid crystal systems. We are interested in
    what information we can gain from a widely-used `out-of-the-box' force field and in future
    work, 
    %MRS: suggest reprhasing as below?XS
    %we will validate whether GAFF is sufficient for reproducing trends in
    %experimental separation properties. 
    we will look at sensitivity of the structural results to force field parameters.
    To clarify our use of GAFF, we have added
    the following text to Secton~\ref{M-method:parameterization} of the Methods on
    page~\pageref{M-method:parameterization}.

    %MRS: suggest pointing out that the LLC's here are not really characteristic of 
    %more commonly studied monomers, so using those LLC force fields isn't really possible.
    New text: ``We chose GAFF because it has been parameterized for use with
    organic molecules. Although researchers have optimized force fields, including
    GAFF, for use with liquid crystals, they are often tuned in order to
    reproduce one or more experimental observables for specific types of liquid 
    crystal molecules.$^{29,30}$ Since liquid crystals exist in a wide range of shapes and
    sizes, including candidate monomers which we may study, we chose to stick with
    a widely used force field. If we can reproduce experimental trends using a 
    single `out-of-the-box' force field, then there will be no need to complicate
    our studies with customized parameter sets."

@@

    We have added the following citations:
    \begin{itemize}
    \item Cook, M. J.; Wilson, M. R. Development of an All-Atom Force Field for the Simulation
    of Liquid Crystal Molecules in Condensed Phases (LCFF). Mol. Cryst. Liq. Cryst.
    2001, 357, 149–165.
    \item Boyd, N. J.; Wilson, M. R. Optimization of the Gaff Force Field to Describe Liquid
    Crystal Molecules: The Path to a Dramatic Improvement in Transition Temperature
    Predictions. Phys. Chem. Chem. Phys. 2015, 17, 24851–24865.
    \end{itemize}

    Since our ultimate goal is to design new monomers for solute-specific
    separations, we will be employing the same techniques used in this work to
    unsynthesized monomers. In most cases, we will not have an experimental dataset
    to use for comparison. Despite all of the structural intricacies which we have
    revealed in this study, one of our main conclusions is that regardless of which
    system we study, 
    %MRS: Not really regardless above, as we only checked systems that were roughly right.
    the radial distribution of monomer components stays
    consistent. From this, we can hypothesize that we will extract similar
    transport results regardless of which configuration we use to conduct
    transport studies. This is an extremely important finding which may have
    been undersold in the original draft of the manuscript. We have added the
    following concluding paragraph to the end of
    Section~\ref{M-section:pore_composition} on
    page~\pageref{M-addition:pore_composition} in order to drive the point home:

@@

    New text: ``Although exactly reproducing the experimental 2D-WAXS pattern
    required attention to minute structural details, we found that all systems
    showed similar radial distribution functions characterized by a gradual radial
    transition from a dense hydrophlic core to hydrophobic tails. This observation
    enables us to study new systems that have not been experimentally characterized
    with the expectation that any additional structural optimization will not
    greatly influence trends in transport property predictions. The compositional
    gradient itself raises questions about the nature of size-exclusion separations
    in systems without well-defined pore boundaries, which potentially could enable
    separations that vary with chemical identity as well as size."

    \item \begin{quote}
    \textit{1. For the simulations performed here, the authors do make a strong
	    case that 5 columns per core gets the best pore-to-pore spacing (Figure 5).
	    However, insufficient time is accessed to see whether the 4-, 5-, and 6-column
	    initial conditions move towards or away from the expected pore spacing over
	    many-microseconds trajectories. That said, for the quenches performed here,
	    5-columns per core seems more plausible than 4.}
    \end{quote}

    Author reply: We agree with the reviewer that 5 columns-per-core is more
    plausible than 4 and that much longer simulations could provide further
    supporting evidence. However, there is very little motion on the hundreds of
    nanoseconds timescale that we simulated, which suggests that we will not gain
    any significant new insight by extending our simulations into the single microsecond
    regime. 
    %MRS: tweaked.
    %It is not feasible to carry out these simulations as it would
    %consume our computational resources with little reward.
    It is not feasible at the present time to push the simulations at least an order of magnitude longer.
    %BJC: The following might apply:
    Some confusion may have been generated by agreement of the disordered basin
    6 column-per-pore system with experimental pore spacing. As mentioned in the
    text on page~\pageref{M-6colperpore}, the vertical stacking distance between 
    monomers in that system is far too large. The columns are `stretched' in the
    $z$-direction allowing the pores take up less space on the $xy$ plane and 
    thus pack closer together.  
    %MRS: for above: Could you propose a textual response to deal with the confusion?

@@

    \item \begin{quote}
    \textit{2. The authors make a pretty compelling case that water causes the
    R-double scattering feature by showing its absence when water is added, but
    that the R-double appears and persists for 200ns after the water makes its way
    into the crystal. However, the authors also show that nonuniformly spaced
    monomers can give rise to the R-Double feature, a seeming counterexample to the
    necessity of water.  Is the R-double feature with water because the monomers
    are now irregularly spaced? What happens when the waters are removed and the
    scattering re-performed in Fig 17c(bottom)?}
    \end{quote}

    Author reply: We thank the reviewer for exposing a potentially unclear
    explanation of the origin of R-double. R-double appears in a dry system with
    non-uniformly spaced monomers \textit{only} if $z$-direction position
    restraints are applied to monomer head groups in order to maintain unevenly
    spaced stacking. As soon as the restraints are released, R-double fades and the
    monomer spacing becomes uniform. We modified the text on page~\pageref{M-modification:rdouble}
    in order to more clearly convey this message:

    Old text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead form pairs that stack less
    than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from neighboring pairs
    of monomers (Figure~\ref{M-fig:staggered_rzplot_norestraints}). Our force field
    causes our system to tend towards uniformly spaced layers. Simulations of
    unevenly spaced systems are only stable if position restraints are applied to
    heavy atoms of the phenyl rings. Additionally, there is little evidence from QM
    studies of stacked $\pi-\pi$ systems that such uneven stacking could be
    energetically stable.$^{52}$%.~\cite{tauer_beyond_2005}"

    Modified text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead are placed in pairs that
    stack less than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from
    neighboring pairs of monomers (Figure~\ref{M-fig:staggered_rzplot_norestraints}).
    Simulations of unevenly spaced systems are only stable while position
    restraints are applied to heavy atoms of the phenyl rings. There is little
    evidence from quantum mechanical studies of stacked $\pi-\pi$ systems that such
    uneven stacking could be energetically stable.$^{52}$ As soon %~\cite{tauer_beyond_2005} As soon
    as we remove position restraints, our system immediately moves towards uniformly spaced
    monomers."

@@

    \item \begin{quote}
    \textit{3. The authors make the case that the R-spots feature comes from the 
    structure of the hexagonal tails, not their tilt. The authors should perform the
    scattering analysis to check if the R-spots are present on the tilted-tails they
    constrain: This quick experiment would nail down their interpretation here, or 
    show that the tilted tails are also plausible (but not observed in this model).}
    \end{quote}

    Author reply: We thank the reviewer for their suggestion. We do not deny
    that the tilt of the monomer tails is a plausible explanation for the
    appearance of reflections that resemble R-spots in many liquid crystal systems.
    While the reviewer's suggested experiment is a good idea, its results in this
    case are ambiguous for a couple reasons. First, the tails may still pack
    hexagonally in the tilted configuration which would independently give rise to
    R-spots. Second, it is necessary to impose high force constant ($>$ 1000 kJ mol$^-1$
    nm$^{-2}$) position restraints in order to maintain a tilt angle close to that
    implied by the 2D WAXS data. The resulting high degree of order in the system 
    gives way to reflections which are easily misinterpretted. We have added the
    following text highlighted in red to the text shown on page~\pageref{M-addition:rspots} in order to
    acknowledge that tail tilt is a valid explanation of R-spot-like reflections:

    %MRS: color is only used twice in this response, and it is not explained at the top 
    % what it is supposed to indicate.  Either that should be clarified and it should be used consistently, 
    or changes should be indicated in other ways.  
    Modified text: ``Previous literature has attributed the R-spots reflection
    in this particular WAXS dataset as the result of tilted alkane
    chains $^{16}$. {\color{red} This explanation is not  % feng_scalable_2014
    unfounded as there are examples in literature of systems where tilted liquid
    crystals give rise to reflections that resemble R-spots.$^{45,46}$} We
    measured the tilt angle of the alkane chains ..."

@@

    \end{itemize}

    \item \begin{quote}
    \textit{These arguments would be stronger if there weren't so many
	    qualifications about simulation timescales and GAFF, in particular its ability
	    to predict the shortest-scale feature (pi-stacking) for this system. How does
	    everything else change if the force-field were tuned to get that feature
	    correct?}
    \end{quote}

    Author reply: We thank the reviewer for their constructive criticism. Many
    of our findings may be relatively independent of the force fields used, but we
    feel that we should include these qualifications in order to not overstate our
    case. It is common for researchers to run simulations for an insufficient
    length of time and to make bold claims. While we feel that we have run our 
    simulations sufficiently long, we still acknowledge the shortcomings of our
    techniques in order to inspire other researchers to build on our initial work.

    We agree with the reviewer that we could tune our force field in order to
    exactly reproduce the $\pi-\pi$ stacking distance seen experimentally. Surely
    there will be some structural differences. There are many ways to tune a force
    field in order to achieve this. As stated earlier, we prefer to work with an
    `out-of-the-box' force field since they are widely used and can be applied 
    %to vast 
    transferably to a large 
    number of systems. The ultimate goal of our work is to create new monomers
    that can achieve specfic separations. If we can gain experimentally-consistent
    insights using our model simulated with GAFF, then we can have increased
    confidence in future predictions generated from systems built with
    unsynthesized monomers.  With that said, we think that it would be interesting
    to repeat our analysis using different and/or optimized force fields in a
    future study in order to comment on the signficance of any differences.

@@

    Old text: ``Future work, based on what has been learned in this study, may
    help further improve the structural agreement between experiment and
    simulation"

    New text: ``Future work, based on what has been learned in this study,
    may help further improve the structural agreement between experiment and
    simulation and test the sensitivity of our current conclusions to a more
    accurate structure."

    \item \begin{quote}
    \textit{The MD-Structure-Factor repository is not available at the provided
	    github link, nor does it appear to be online anywhere. So, the correctness of
	    that code cannot be checked.}
    \end{quote}

    Author reply: We thank the reviewer for pointing this out and apologize for
    our mistake. The referenced GitHub repository was set to private due to
    miscommunication with our collaborators but it has now been made public. 
    %MRS: not needed.
    %Please feel free to read through the code. 
    We have performed checks to ensure that the
    %MRS: can you be more specific what is meant by ``works properly''?
    code works properly, especially when simulating monoclinic unit cells. For
    example, we generated test trajectories containing thousands of frames made up
    of configurations of randomly placed point scatterers in a monoclinic unit cell.
    %MRS: tweak below.
    %We expect that the simulated diffraction pattern will have a Gaussian shape.
    In this case, the simulated diffraction pattern will be Gaussian.
    This exact result is shown in the Figure~\ref{S-fig:xrd_noise} of the SI.
    With improper treatment of the monoclinic geometry, the gaussian shape is
    skewed into an elliptical shape.

@@

    Revised text: ``The calculated value of ionic conductivity is 5 times higher than
    experiment likely because we simulated infinitely long, aligned pores {\color{red} and
    because our model over-predicts the diffusivity of sodium}. The ionic conductivity 
    measurement to which we are comparing was done with an \SI{80}{\micro\metre}-thick film, 
    nearly 10,000 times thicker than our simulated system. The thick film is likely 
    imperfectly aligned and has defects leading to non-contiguous pores. It has been shown
    that there is a large dependence of ionic conductivity on the alignment of the pores.
    The ionic conductivity of an isotropically aligned film is ca. 85 times lower than that
    of the nearly aligned film to which we are comparing$^{16}$. 
    {\color{red} Additionally, sodium ions parameterized with AMBER parameters exhibit bulk
    water diffusion coefficients that are two times greater than experiment.$^{56}$ We cannot 
    definitively say how this ratio changes in the confined LLC pore environment, but
    it is likely that sodium diffuses faster in our system which contributes to a higher 
    ionic conductivity than experiment.}"

	% Amber parameters used for sodium
	% most studies with conductivity look at potassium
	% look at diffusivities

	\item \begin{quote}
    \textit{If the authors either (a) strengthen the cases for their proposed structural
    insights being  highly likely (not just plausible), or (b) make a stronger case for
    why these particular insights have important ramifications for our understanding of 
    LLC's (their significance), then it would be easier to publish this immediately. This
    reviewer's opinion is that this work would find a broader interested audience with a more
    concise focus on these physical insights, which could be accomplished by moving more
    content (e.g., dihedral correlations) to the SI.}
    \end{quote}

    Author reply: We feel that the additions and modifications to the sections cited above 
    have made the intent of our work clearer and helped communicate the significance of 
    our learned insights for future work on this topic. We have clarified the purpose of our work
    and what we expect to learn from it. We have commented on the significance of knowing the
    number of columns-per-pore. We have emphasized the observation that the radial distribution
    of monomer components is relatively independent of the system's initial configuration
    which is important for future high-throughput work. We have clarified and strengthened our
    reasoning for the existence of the R-double feature.

    We sympathize with the reviewer's desire for a more concise manuscript but feel that most 
    of the information presented is necessary in order to come to a full understanding of the 
    work we've performed (an opinion shared by Reviewer 2). 
    %MRS: tweak
    %As a compromise, we've made efforts to shorten the manuscript by 
    We have attempted to satisfy both viewpoints, by carefully examining the manuscript for figures 
    and discussions which draw the reader's attention from the important physical insights, and moving 
    these to the SI. 
    %MRS: could you clarify this?  Maybe just state how many pages were moved. 
    Excluding additional citations (new citations
    take up about a page in the new draft), we have reduced the total length of the manuscript 
    by 3 pages (including any additions documented above). 
    %MRS: tweak
    %The following changes made this possible:
    The following sections were moved to the SI, with only the key information left in the main paper:
    \begin{itemize}

### 2018-11-21 Ben_Manuscripts/structure_paper/Final_Draft.tex
Final reviewer response

  % BJC: reworked text
  \label{rework:purpose}
  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, will enhance our comprehension of the role
  of monomer structure in solute transport and provide guidelines to reduce the
  large chemical space available to design monomers for creation of
  separation-specific membranes. It is important that we model the system with a
  nanoscopic pore structure that is maximally consistent with experimental data
  in order to approximate the chemical environment experienced by solutes in 
  future transport studies. Atomistic MD simulations can provide the required
  level of detail (Figure \ref{fig:detailed_pore}), assuming the force fields are
  sufficiently accurate. With such an atomistic model, we can directly observe
  molecular-level solute transport and suggest governing mechanisms. We can also
  observe how the choice of head group interacts with solutes of interest. In
  addition, we can interchange counterions which may influence both the pore size
  and the strength of the Donnan potential.

@@

  We parameterized the interaction potential for the liquid crystal monomer
  Na-GA3C11 using the Generalized AMBER Force Field (GAFF)
  \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. We chose GAFF because it has been parameterized
  for use with organic molecules. Although researchers have optimized force
  fields, including GAFF, for use with liquid crystals, they are often tuned in 
  order to reproduce one or more experimental observables for specific types of
  liquid crystal molecules~\cite{cook_development_2001,boyd_optimization_2015}.
  Since liquid crystals exist in a wide range of shapes and sizes, including candidate
  monomers which we may study, we chose to stick with a widely used force field.
  If we can reproduce experimental trends using a single `out-of-the-box' force
  field, then there will be no need to complicate our studies with customized
  parameter sets. We assigned atomic charges using the am1bccsym method of
  \texttt{molcharge} shipped with QUACPAC from Openeye Scientific Software. We
  ran all molecular dynamics simulations using GROMACS 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the
  equilibrium configuration is sensitive to some while insensitive to others. For
  example, we find the starting pore radius, defined as the distance of a chosen
  head group carbon from the pore's central axis, does not influence the
  equilibrium structure if one chooses a reasonable value. The initial distance
  between pores, within a wide range, also has little effect on the equilibrated
  structure. However, one should not start them too close or there will be high
  energy repulsions during early equilibration. We chose a pore radius of 0.5 nm
  and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental
  value of 4.12 nm for our initial configurations. A sensitivity analysis of both
  parameters is presented in the SI, Section
  \ref{S-section:initial_config_dependence}. The distance between vertically
  stacked monomers, the $xy$ position of monomers with respect to vertically
  adjacent monomers, and the number of columns per pore do influence the
  equilibrated structure and require further justification for their choices. We
  rely on experimental data to inform them.

@@

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns per pore. See Figure 
  \ref{S-fig:stacking} of the SI for a detailed illustration 
  of the initial configurations in each mode.

  The number of columns per pore is unknown, as stated in Question
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of columns per pore. We built systems in the sandwiched and parallel
  displaced configurations with 4, 5, 6, 7 and 8 columns per pore.

@@

  Our simulations best support a model built with 5 monomer columns per pore
  based on the measured equilibrated pore-to-pore distances. To identify the
  density of monomer columns around each pore, addressing Question
  \ref{point:monomernum} in the introduction, we ran simulations of systems
  created with 4--8 columns per pore. We built systems in both the parallel
  displaced and sandwiched configurations and equilibrated them according to the
  dry equilibration procedure. We tested all systems with an initial vertical
  monomer spacing, $\mathit{d}$, of 3.7 \AA~in accordance with R-$\pi$. We tested
  4 additional systems with monomers initially spaced 5 \AA~apart vertically (see
  Section \ref{S-section:initial_pore_spacing} of the SI for more details on
  sensitivity to initial monomer spacing). We considered the pore-to-pore
  spacing to be equilibrated as defined in Section~\ref{method:equil_time}.
  Figure~\ref{fig:p2p} shows the equilibrated pore-to-pore distances for all
  systems tested.

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=\linewidth]{p2p.pdf}
	\caption{Systems with 5 columns per pore have equilibrated pore spacings closest to
			 the experimental value of 4.12 nm. The equilibrated pore spacing of the model 	
			 increases as the number of columns in each pore increases.}~\label{fig:p2p}
  \end{figure}

@@

  \label{6colperpore}
  Systems built with 5 columns around each pore equilibrate to a pore spacing
  that is most consistent with the experimental value of 4.12 nm derived from
  SAXS measurements (Figure~\ref{fig:SAXS}). Ordered basin systems built with 4
  columns per pore equilibrate to an average pore spacing 0.25 nm lower than
  experiment. Ordered basin systems built with 6 columns per pore, have an
  equilibrated pore spacing ca. 0.50 nm higher than experiment. Monomers in
  disordered basin systems built with 6 columns-per-pore agree with experimental
  pore-to-pore distances within error, but stack too far apart. 6
  column per pore disordered sandwiched and disordered parallel displaced
  configurations stack $\sim$~4.87 and 4.94 \AA~apart respectively, which is
  $\sim$ 1.2~\AA~further apart than suggested by experiment. 5 column-per-pore
  systems stack, at a maximum, 0.9 \AA~further apart than experiment (see
  Table~\ref{table:correlation_length}). The remainder of this discussion will
  therefore focus on the analysis of systems built with 5 columns per pore.

  %BJC: added.
  \label{addition:pore_radius}
  The number of columns per pore dictates the density of interaction sites within
  each pore and the pore radius. A higher number of interaction sites will 
  surely play a role in transport of molecules that have an affinity for the monomer
  head groups. The pore radius increases with the number of 
  columns per pore (See Figure~\ref{S-fig:pore_radius}). This radius is likely 
  related to the size of the molecules which can be excluded by this type of 
  membrane. One may be able to control LLC membrane pore size by selecting monomers
  that are stable in configurations with more columns-per-pore.

@@

  We can better understand the structure of the system by simulating XRD
  patterns produced from equilibrated MD trajectories (see
  Section~\ref{method:xrd}) and comparing these simulated patterns to experiment,
  addressing Question~\ref{point:xrdmatch} in the Introduction. We tested systems
  built with 5 columns per pore in the parallel displaced and sandwiched
  configurations at 300 K in the ordered and disordered basins, as those were the
  most consistent with the most unambiguous XRD features, the pore-to-pore
  spacing and the stacking distance, as described in
  Section~\ref{section:mon_per_pore}.  We generated simulated patterns using the
  locally equilibrated portion of each simulation trajectory (see
  Section~\ref{method:equil_time}). There are two important factors to take into
  account when looking at differences between experimental and simulated XRD
  patterns.  First, the simulated diffraction patterns have some noise,
  especially along the $q_z$ axis at $q_r$=0, which is where several of the more
  interesting features are located. This is due to the angle averaging of the 3D
  structure factor around the $q_z$-axis, meaning there are fewer samples as
  $q_r$ approaches 0.  More importantly, our simulations do not appear to be long
  enough to sample truly independent configurations within their respective
  metastable basins (see Section~\ref{section:slow_dynamics}), meaning we must
  take care in interpreting the results of the XRD. The amount of independent
  configurations required to converge along the $q_z$ axis is analyzed in SI,
  Section \ref{S-section:xrd_noise}. The simulated patterns generated for all
  systems studied are shown and compared to experiment in
  Figure~\ref{fig:XRDsim}.

>>>>

  % BJC: reworked text
  \label{rework:purpose}
  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, will enhance our understanding of the role
  of monomer structure in solute transport and provide guidelines to reduce the 
  large chemical space available to design monomers for creation of 
  separation-specific membranes. It is important that we model the system with
  a nanoscopic pore structure that is maximally consistent with experimental data
  in order to best approximate the chemical environment experienced by solutes
  in future simulated transport studies. Atomistic MD simulations can provide the required
  level of detail (Figure \ref{fig:detailed_pore}), assuming the force fields are
  sufficiently accurate. With such an atomistic model, we can directly observe
  molecular-level solute transport and suggest governing mechanisms. We can also
  observe how the choice of head group interacts with solutes of interest. In
  addition, we can interchange counterions which may influence both the pore size
  and the strength of the Donnan potential.

@@

  We parameterized the interaction potential for the liquid crystal monomer
  Na-GA3C11 using the Generalized AMBER Force Field (GAFF)
  \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. We chose GAFF because it has been parameterized
  for use with organic molecules. Although researchers have optimized force
  fields, including GAFF, for use with liquid crystals, they are often tuned in 
  order to reproduce one or more experimental observables for specific types of
  liquid crystal molecules~\cite{cook_development_2001,boyd_optimization_2015}.
  Na-GA3C11 does not have a structure characteristic of more commonly studied 
  liquid crystal monomers so we chose to stick with a widely used force field.
  If we can reproduce experimental trends using a single `out-of-the-box' force
  field, then there will be no need to complicate our studies with customized
  parameter sets. We assigned atomic charges using the am1bccsym method of
  \texttt{molcharge} shipped with QUACPAC from Openeye Scientific Software. We
  ran all molecular dynamics simulations using GROMACS 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the
  equilibrium configuration is sensitive to some while insensitive to others. For
  example, we find the starting pore radius, defined as the distance of a chosen
  head group carbon from the pore's central axis, does not influence the
  equilibrium structure if one chooses a reasonable value. The initial distance
  between pores, within a wide range, also has little effect on the equilibrated
  structure. However, one should not start them too close or there will be high
  energy repulsions during early equilibration. We chose a pore radius of 0.5 nm
  and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental
  value of 4.12 nm for our initial configurations. A sensitivity analysis of both
  parameters is presented in the SI, Section
  \ref{S-section:initial_config_dependence}. The distance between vertically
  stacked monomers, the $xy$ position of monomers with respect to vertically
  adjacent monomers, and the number of columns-per-pore do influence the
  equilibrated structure and require further justification for their choices. We
  rely on experimental data to inform them.

@@

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns-per-pore. See Figure 
  \ref{S-fig:stacking} of the SI for a detailed illustration 
  of the initial configurations in each mode.

  The number of columns-per-pore is unknown, as stated in Question
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of columns-per-pore. We built systems in the sandwiched and parallel
  displaced configurations with 4, 5, 6, 7 and 8 columns-per-pore.

@@

  Our simulations best support a model built with 5 monomer columns-per-pore
  based on the measured equilibrated pore-to-pore distances. To identify the
  density of monomer columns around each pore, addressing Question
  \ref{point:monomernum} in the introduction, we ran simulations of systems
  created with 4--8 columns-per-pore. We built systems in both the parallel
  displaced and sandwiched configurations and equilibrated them according to the
  dry equilibration procedure. We tested all systems with an initial vertical
  monomer spacing, $\mathit{d}$, of 3.7 \AA~in accordance with R-$\pi$. We tested
  4 additional systems with monomers initially spaced 5 \AA~apart vertically (see
  Section \ref{S-section:initial_pore_spacing} of the SI for more details on
  sensitivity to initial monomer spacing). We considered the pore-to-pore
  spacing to be equilibrated as defined in Section~\ref{method:equil_time}.
  Figure~\ref{fig:p2p} shows the equilibrated pore-to-pore distances for all
  systems tested.

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=\linewidth]{p2p.pdf}
	\caption{Systems with 5 columns-per-pore have equilibrated pore spacings closest to
			 the experimental value of 4.12 nm. The equilibrated pore spacing of the model 	
			 increases as the number of columns in each pore increases.}~\label{fig:p2p}
  \end{figure}

@@

  \label{6colperpore}
  Systems built with 5 columns around each pore equilibrate to a pore spacing
  that is most consistent with the experimental value of 4.12 nm derived from
  SAXS measurements (Figure~\ref{fig:SAXS}). Ordered basin systems built with 4
  columns-per-pore equilibrate to an average pore spacing 0.25 nm lower than
  experiment. Ordered basin systems built with 6 columns-per-pore, have an
  equilibrated pore spacing ca. 0.50 nm higher than experiment. Monomers in 
  disordered basins systems stack too far apart, therefore they are not 
  suitable candidate structures. Although their pore-to-pore spacing is in
  close agreement with experiment, the 6 column-per-pore disordered 
  sandwiched and disordered parallel displaced configurations stack 
  $\sim$~4.87 and 4.94 \AA~apart respectively, which is $\sim$ 
  1.2~\AA~further apart than suggested by experiment. The monomer columns in
  these systems are effectively stretched in the $z$-direction which allows
  them and the pores to pack closer together on the $xy$ plane. 5 column-per-pore
  systems stack, at a maximum, 0.9 \AA~further apart than experiment 
  (see Table~\ref{table:correlation_length}). The remainder of this discussion
  will therefore focus on the analysis of systems built with 5 columns-per-pore.

  %BJC: added.
  \label{addition:pore_radius}
  The number of columns-per-pore dictates the density of interaction sites within
  each pore and the pore radius. A higher number of interaction sites will 
  surely play a role in transport of molecules that have an affinity for the monomer
  head groups. The pore radius increases with the number of 
  columns-per-pore (See Figure~\ref{S-fig:pore_radius}). This radius is likely 
  related to the size of the molecules which can be excluded by this type of 
  membrane. One may be able to control LLC membrane pore size by selecting monomers
  that are stable in configurations with more columns-per-pore.

  \begin{table}[h]
  \centering
  \begin{tabular}{cccc}
  \toprule
  System             & $\mathit{d}$ (\AA) & $\mathit{d}_{equil}$ (\AA) & Correlation Length (\AA) \\
  \midrule
  Sandwiched         & 3.7                &    4.27 $\pm$ 0.03         & 4.2 $\pm$ 0.8            \\
  Parallel Displaced & 3.7                &    4.33 $\pm$ 0.04         & 14.5 $\pm$ 1.3           \\ 
  Sandwiched         & 5.0                &    4.48 $\pm$ 0.07         & 3.2 $\pm$ 0.9            \\
  Parallel Displaced & 5.0                &    4.60 $\pm$ 0.08         & $<$ $d_{equil}$ \\ 
  Experiment         & --                 &    3.70                    & 9 $\pm$ 1               \\
  \bottomrule
  \end{tabular}
  \caption{For 5 column-per-pore systems, the equilibrated vertical stacking distance, 
  $\mathit{d}_{equil}$, of monomers in the disordered basin is approximately 0.6 \AA higher than
  their initial configuration stacking distance, $\mathit{d}$, while $\mathit{d}_{equil}$ is 
  smaller than $\mathit{d}$ for disordered basin systems, but 0.9 \AA higher than experiment. 
  The correlation length is larger for systems in the ordered basin.}
  \label{table:correlation_length}
  \end{table}

@@

  We can better understand the structure of the system by simulating XRD
  patterns produced from equilibrated MD trajectories (see
  Section~\ref{method:xrd}) and comparing these simulated patterns to experiment,
  addressing Question~\ref{point:xrdmatch} in the Introduction. We tested systems
  built with 5 columns-per-pore in the parallel displaced and sandwiched
  configurations at 300 K in the ordered and disordered basins, as those were the
  most consistent with the most unambiguous XRD features, the pore-to-pore
  spacing and the stacking distance, as described in
  Section~\ref{section:mon_per_pore}.  We generated simulated patterns using the
  locally equilibrated portion of each simulation trajectory (see
  Section~\ref{method:equil_time}). There are two important factors to take into
  account when looking at differences between experimental and simulated XRD
  patterns.  First, the simulated diffraction patterns have some noise,
  especially along the $q_z$ axis at $q_r$=0, which is where several of the more
  interesting features are located. This is due to the angle averaging of the 3D
  structure factor around the $q_z$-axis, meaning there are fewer samples as
  $q_r$ approaches 0.  More importantly, our simulations do not appear to be long
  enough to sample truly independent configurations within their respective
  metastable basins (see Section~\ref{section:slow_dynamics}), meaning we must
  take care in interpreting the results of the XRD. The amount of independent
  configurations required to converge along the $q_z$ axis is analyzed in SI,
  Section \ref{S-section:xrd_noise}. The simulated patterns generated for all
  systems studied are shown and compared to experiment in
  Figure~\ref{fig:XRDsim}.
